### 2020-07-28 manuscript_code_20200722.Rmd
blight added

---
title: "Code for Breast Cancer Pain Study"
output:
  html_document:
    toc: true
    toc_depth: 2
    number_sections: true
    fig_width: 12
    fig_height: 6
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
library(knitr)
opts_chunk$set(echo=TRUE,autodeep=TRUE,cache.comments=FALSE,warning=FALSE,message=FALSE,cache.lazy=FALSE)
```

# Data pre-processing

```{r,echo=FALSE}
library(openxlsx)
library(lattice)
library(quantreg)
library(ggplot2)
library(pscl)
library(boot)
library(MASS)
library(sjPlot)
library(nonnest2)
library(vcd)
library(AER)
library(faraway)
library(tidyverse)
library(table1)
library(RColorBrewer)
library(ggpubr)

transform.p = function(x,digits=3) {
  y = as.character(round(x,digits=digits))
  y[which(y==0)] = "<.001"
  return(y)
}

```

```{r,echo=FALSE}
params = as.list(NULL)
params$days.cutoff = 12
params$bin.cutoff = 0
params$race = "All"
params$occupation = FALSE
params$patientsData = "PatientsData_merged_2020-07-15.xlsx"
params$PainData = "PCM_Dates_merged_2020-06-29.xlsx"

days.cutoff = as.numeric(params$days.cutoff)*(365/12)
bin.cutoff = as.numeric(params$bin.cutoff) # from what number, pain considered as "no pain"?
export.data = FALSE # save merged data as excel format?

## Load data
patients_file=read.xlsx(xlsxFile=as.character(params$patientsData),sheet=1,detectDates=TRUE)
pcm_file=read.xlsx(xlsxFile=as.character(params$PainData),sheet=1,detectDates=TRUE)

colnames(patients_file)[which(colnames(patients_file)=="POVERTY")] = "Poverty"
colnames(patients_file)[which(colnames(patients_file)=="VACANCY")] = "Vacancy"
colnames(patients_file)[which(colnames(patients_file)=="WALKABILITY")] = "Walkability"
colnames(patients_file)[which(colnames(patients_file)=="UNEMPLOYMENT")] = "Unemployment"
colnames(pcm_file)[which(colnames(pcm_file)=="a_CCMA06")] = "Pain"
```

**_Patients selection_**.

* Cleaned datasets with new dx_date (initial diagnosis date) was released - R21_Sample_wDxDate.     
* Only TN considered. Other states were excluded. (3836 $\rightarrow$ 2628)      
* Some duplicated samples were removed.   
    - There are 43 duplicated profile keys in the merged data set.     
    - e.g. profile key==90006  
* There are a lot of missing values in the important variables such as Marital status, Occupation, Primary Payer, POVERTY, and etc. The missing values are systematic rather than just random, for example, if one patient is missing in one variable then it is also missing in the other variables. Therefore, we will only include the patients with no missing values in any of the important covariates.

* missing, NA -> Unknown

**_Pain data (PCM) selection_**.

* Pain variable (a_CCMA06) observed after at least 60 days prior to dx_date only considered.   
* Pain variable (a_CCMA06) within 365 days after the initial diagnosis date (dx_date) has been chosen.     
* Mean was taken for Pain by patients.

**_Poverty variable collection_**.

* "joined_zipcode" file merged to get data about poverty, vacancy, unemployment levels for more zip codes.   
* Poverty data (from "joined_zipcode")

    - For poverty variable, group mean used for each zip code.   
    - Only 43 zip codes are documented in poverty file. On the other hand, there are 165 zip codes (TN) in Patients file. This produced many missing values for the poverty variable.

**_Further data selection_**.

* Race = "Other" or "Unknown" removed. (171 Samples excluded)
* NA's in Pain variable removed. (247 samples excluded)

```{r,echo=FALSE}
pcm_profile_key=sort(unique(pcm_file$profile_key))
patients_profile_key=sort(unique(patients_file$profile_key))

cat(paste("Number of patients in the cohort (PatientsData) =",length(patients_profile_key)),"\n")
pcm_subset = pcm_file[which((abs(pcm_file$dx_diff_days)<days.cutoff) & (pcm_file$dx_diff_days>(-60))),]

pcm_new = cbind(aggregate(. ~ profile_key, pcm_subset[,c(1,3)], FUN=function(x){round(mean(x))}),
                pcm_Q1st = aggregate(. ~ profile_key, pcm_subset[,c(1,3)], FUN=function(x){round(quantile(x)[2])})[,2],
                pcm_Q2nd = aggregate(. ~ profile_key, pcm_subset[,c(1,3)], FUN=function(x){round(quantile(x)[3])})[,2],
                pcm_Q3rd = aggregate(. ~ profile_key, pcm_subset[,c(1,3)], FUN=function(x){round(quantile(x)[4])})[,2],
                pcm_min = aggregate(. ~ profile_key, pcm_subset[,c(1,3)], FUN=function(x){min(x)})[,2],
                pcm_max = aggregate(. ~ profile_key, pcm_subset[,c(1,3)], FUN=function(x){max(x)})[,2],
                pcm_bin = aggregate(. ~ profile_key, pcm_subset[,c(1,3)], FUN=function(x){as.numeric((min(x)>bin.cutoff))})[,2])
colnames(pcm_new) = c(colnames(pcm_new)[1],paste(c("mean","Q1st","Q2nd","Q3rd","min","max","bin"),colnames(pcm_new)[2],sep="."))

pcm_count = aggregate(. ~ profile_key, pcm_subset[,c(1,3)], FUN=function(x){length(x)})
colnames(pcm_count)[2] = "count.Pain"
pcm_var = merge(x=pcm_new,y=pcm_count,by="profile_key",all=T)
profile_key = pcm_var$profile_key

## Add chemo variable: afchemo / bfchemo (after/before chemo)
afchemo = bfchemo = matrix(NA,ncol=ncol(pcm_var),nrow=length(profile_key))
afchemo[,1] = bfchemo[,1] = profile_key
colnames(bfchemo) = paste("bfchemo",colnames(pcm_var),sep=".")
colnames(afchemo) = paste("afchemo",colnames(pcm_var),sep=".")
colnames(bfchemo)[1] = colnames(pcm_var)[1]
colnames(afchemo)[1] = colnames(pcm_var)[1]

for (i in 1:length(profile_key)) {
  x = profile_key[i]
  temp_rows=which(pcm_subset$profile_key==x)
  pcm_subset[temp_rows,]

  diff.date = (pcm_subset$dx_diff_days[temp_rows] - pcm_subset$chemo_diff_days[temp_rows])[1] # chemo.date - dx.date
  if (is.na(diff.date)) {
    next
  } else if ((diff.date>(-365)) & (diff.date<(days.cutoff))) {
    bfchemo.table = pcm_subset[temp_rows[which(pcm_subset$chemo_diff_days[temp_rows]<=0)],]
    afchemo.table = pcm_subset[temp_rows[which(pcm_subset$chemo_diff_days[temp_rows]>0)],]

    if (dim(bfchemo.table)[1]>0) {
      bfchemo[i,2:ncol(bfchemo)] = c(round(mean(bfchemo.table$Pain)),
                                     round(quantile(bfchemo.table$Pain)[2]),
                                     round(quantile(bfchemo.table$Pain)[3]),
                                     round(quantile(bfchemo.table$Pain)[4]),
                                     min(bfchemo.table$Pain),
                                     max(bfchemo.table$Pain),
                                     as.numeric((min(bfchemo.table$Pain)>bin.cutoff)),
                                     dim(bfchemo.table)[1])
    }
    if (dim(afchemo.table)[1]>0) {
      afchemo[i,2:ncol(afchemo)] = c(round(mean(afchemo.table$Pain)),
                                     round(quantile(afchemo.table$Pain)[2]),
                                     round(quantile(afchemo.table$Pain)[3]),
                                     round(quantile(afchemo.table$Pain)[4]),
                                     min(afchemo.table$Pain),
                                     max(afchemo.table$Pain),
                                     as.numeric((min(afchemo.table$Pain)>bin.cutoff)),
                                     dim(afchemo.table)[1])
    }
  }
  # cat(paste(i),"\n")
}
pcm_var2 = merge(x=pcm_var,y=bfchemo,by="profile_key",all=T)
pcm_var2 = merge(x=pcm_var2,y=afchemo,by="profile_key",all=T)
# head(pcm_var2,n=20)

## Add surgery variable: afsurgery / bfsurgery (after/before surgery)
afsurgery = bfsurgery = matrix(NA,ncol=ncol(pcm_var),nrow=length(profile_key))
afsurgery[,1] = bfsurgery[,1] = profile_key
colnames(bfsurgery) = paste("bfsurgery",colnames(pcm_var),sep=".")
colnames(afsurgery) = paste("afsurgery",colnames(pcm_var),sep=".")
colnames(bfsurgery)[1] = colnames(pcm_var)[1]
colnames(afsurgery)[1] = colnames(pcm_var)[1]

for (i in 1:length(profile_key)) {
  x = profile_key[i]
  temp_rows=which(pcm_subset$profile_key==x)
  pcm_subset[temp_rows,]

  diff.date = (pcm_subset$dx_diff_days[temp_rows] - pcm_subset$surgery_diff_days[temp_rows])[1] # surgery.date - dx.date
  if (is.na(diff.date)) {
    next
  } else if ((diff.date>(-365)) & (diff.date<(days.cutoff))) {
    bfsurgery.table = pcm_subset[temp_rows[which(pcm_subset$surgery_diff_days[temp_rows]<=0)],]
    afsurgery.table = pcm_subset[temp_rows[which(pcm_subset$surgery_diff_days[temp_rows]>0)],]

    if (dim(bfsurgery.table)[1]>0) {
      bfsurgery[i,2:ncol(bfsurgery)] = c(round(mean(bfsurgery.table$Pain)),
                                     round(quantile(bfsurgery.table$Pain)[2]),
                                     round(quantile(bfsurgery.table$Pain)[3]),
                                     round(quantile(bfsurgery.table$Pain)[4]),
                                     min(bfsurgery.table$Pain),
                                     max(bfsurgery.table$Pain),
                                     as.numeric((min(bfsurgery.table$Pain)>bin.cutoff)),
                                     dim(bfsurgery.table)[1])
    }
    if (dim(afsurgery.table)[1]>0) {
      afsurgery[i,2:ncol(afsurgery)] = c(round(mean(afsurgery.table$Pain)),
                                     round(quantile(afsurgery.table$Pain)[2]),
                                     round(quantile(afsurgery.table$Pain)[3]),
                                     round(quantile(afsurgery.table$Pain)[4]),
                                     min(afsurgery.table$Pain),
                                     max(afsurgery.table$Pain),
                                     as.numeric((min(afsurgery.table$Pain)>bin.cutoff)),
                                     dim(afsurgery.table)[1])
    }
  }
  # cat(paste(i),"\n")
}

pcm_var2 = merge(x=pcm_var2,y=bfsurgery,by="profile_key",all=T)
pcm_var2 = merge(x=pcm_var2,y=afsurgery,by="profile_key",all=T)
# head(pcm_var2,n=20)

dmd = merge(pcm_var2,patients_file,by="profile_key",all=T)
# if (export.data) {
#   write.xlsx(x=dmd,file=paste0(dataDir,"PCM_PatientsVar_merged_DX",days.cutoff,"daysRule_",today,".xlsx"))
# }

```

```{r,echo=FALSE}
## Define factor variables:
# table(dmd$Tobacco.History)
dmd$Tobacco.History = factor(dmd$Tobacco.History,levels=c("Never","Smoked","Unknown"))

# table(dmd$Histology)
dmd$Histology = factor(dmd$Histology,levels=c("ductal","lobular","Other","Unknown"))

# table(dmd$Recurrence)
dmd$Recurrence = factor(dmd$Recurrence,levels=c("disease-free","Recurrence","Unknown"))

# table(dmd$Primary.Site.1)
dmd$Primary.Site.1 = factor(dmd$Primary.Site.1,levels=c("Lower","Central","Upper","Unknown"))

# table(dmd$Primary.Site.2)
dmd$Primary.Site.2 = factor(dmd$Primary.Site.2,levels=c("Inner","Midline","Outer","Overlapping","Unknown"))

# table(dmd$HER2)
dmd$HER2[(dmd$HER2 %in% c("Borderline","Test.not.done","Unknown"))] = "Unknown"
dmd$HER2 = factor(dmd$HER2,levels=c("Negative.Normal","Positive.Elevated","Unknown"))

# table(dmd$ER)
dmd$ER[(dmd$ER %in% c("Test.not.done","Unknown"))] = "Unknown"
dmd$ER = factor(dmd$ER,levels=c("Negative.Normal","Positive.Elevated","Unknown"))

# table(dmd$PR)
dmd$PR[(dmd$PR %in% c("Test.not.done","Unknown"))] = "Unknown"
dmd$PR = factor(dmd$PR,levels=c("Negative.Normal","Positive.Elevated","Unknown"))

# table(dmd$Metastasis)
dmd$Metastasis = factor(dmd$Metastasis,levels=c("No","Yes"))

# Relevel factor variables
dmd$race.f = relevel(factor(dmd$race.f),"White")
# table(dmd0$race.f)

dmd$marital.status.f = relevel(factor(dmd$marital.status.f),"Married")
dmd$marital.status.f = factor(dmd$marital.status.f, levels=c("Married","Single","Divorced","Widowed","Unknown"))
# table(dmd$marital.status.f)

dmd$Usual.Occupation.f = relevel(factor(dmd$Usual.Occupation.f),"Employed")
# table(dmd0$Usual.Occupation.f)

dmd$Primary.Payer.f = relevel(factor(dmd$Primary.Payer.f),"Insurance")
# table(dmd0$Primary.Payer.f)

# table(dmd$race.f)

## Usual Occupation
occupation = dmd$Usual.Occupation.f
occupation[which(occupation=="Disabled")] = "Unemployed"
if (params$occupation) {
  occupation[which(occupation=="Retired")] = "Unemployed"
}
occupation.f = factor(occupation)
dmd$Usual.Occupation.f2 = occupation.f
# table(dmd$Usual.Occupation.f2)

## Primary Payer
insurance = dmd$Primary.Payer.f
insurance[which(insurance=="Private")] = "Insurance"
insurance.f = factor(insurance)
dmd$Primary.Payer.f2 = insurance.f
# table(dmd$Primary.Payer.f2)

## Marital Status
marriage = dmd$marital.status.f
marriage[which(marriage=="Widowed")] = "Single"
marriage.f = factor(marriage)
dmd$marital.status.f2 = marriage.f
# table(dmd$marital.status.f2)

## Poverty -> factor (low/high)
Poverty = rep("low",dim(dmd)[1])
Poverty[which(dmd$Poverty>median(dmd$Poverty,na.rm=T))] = "high"
Poverty.f = factor(Poverty)
Poverty.f = relevel(Poverty.f,"low")
dmd$Poverty.f = Poverty.f
# table(dmd$Poverty.f)

## Stage
dmd$stage.f = factor(dmd$stage.f)
dmd$stage.f = relevel(dmd$stage.f,"1")

## Age
age = rep("Low",dim(dmd)[1])
age[which(dmd$age.f==1)] = "High"
dmd$age.f = factor(age)
dmd$age.f = relevel(dmd$age.f,"Low")

## Chemo -> factor (0=no, 1=yes)
chemo = rep("Untreated",dim(dmd)[1])
chemo[which(dmd$afchemo.count.Pain>0)] = "Treated"
chemo.f = factor(chemo)
chemo.f = relevel(chemo.f,"Treated")
dmd$chemo.f = chemo.f
# table(dmd$chemo.f)

## surgery -> factor (0=no, 1=yes)
surgery = rep("Untreated",dim(dmd)[1])
surgery[which(dmd$afsurgery.count.Pain>0)] = "Treated"
surgery.f = factor(surgery)
surgery.f = relevel(surgery.f,"Treated")
dmd$surgery.f = surgery.f
# table(dmd$surgery.f)

if (! params$race=="All") {
  dmd = dmd[which(dmd$race.f==params$race),]
}

# unknownRaceN = length(which(dmd$race.f %in% c("Other","Unknown")))
# dmd = dmd[-which(dmd$race.f %in% c("Other","Unknown")),]
dmd$race.f[which(dmd$race.f %in% c("Other","Unknown"))] = "Unknown"
# cat(paste("Unknown and Other in Race removed (",unknownRaceN,"), Black and White remained N =",dim(dmd)[1]),"\n")
dmd$race.f = factor(dmd$race.f)

## Re-name variables
# colnames(cdmd.pvt)[which(grepl(pattern="surgery",colnames(cdmd.pvt))==T)]
colnames(dmd)[which(colnames(dmd)=="Usual.Occupation.f2")] = "Occupation"
colnames(dmd)[which(colnames(dmd)=="marital.status.f2")] = "Marital.status"
colnames(dmd)[which(colnames(dmd)=="Primary.Payer.f2")] = "Insurance.type"
colnames(dmd)[which(colnames(dmd)=="race.f")] = "Race"
colnames(dmd)[which(colnames(dmd)=="stage.f")] = "Cancer.stage"
colnames(dmd)[which(colnames(dmd)=="age.f")] = "Age"
colnames(dmd)[which(colnames(dmd)=="chemo.f")] = "Chemo"
colnames(dmd)[which(colnames(dmd)=="surgery.f")] = "Surgery"
colnames(dmd)[which(colnames(dmd)=="Tobacco.History")] = "Tobacco.history"
```

```{r,echo=FALSE}
## Select samples with Poverty values
dmd.pvt = dmd[which(!is.na(dmd$Poverty)==T),]
zipcode.pvt = sort(unique(dmd.pvt$zip))
cat(paste("Number of patients with Poverty reported =",dim(dmd.pvt)[1]),"\n")

dmd.pvt = dmd.pvt[-which(dmd.pvt$Poverty>70),]
dmd.pvt = dmd.pvt[-which((dmd.pvt$Vacancy>40) & (dmd.pvt$mean.Pain<2)),]
cat(paste("Three outliers excluded"),"\n")

cdmd.pvt = dmd.pvt
colnames(cdmd.pvt)[which(grepl(pattern="mean",colnames(cdmd.pvt))==T)] = c("Pain","Pain.bfchemo","Pain.afchemo","Pain.bfsurgery","Pain.afsurgery")
NANs = length(which(is.na(cdmd.pvt$Pain)))
cdmd.pvt = cdmd.pvt[which(!is.na(cdmd.pvt$Pain)),]
cat(paste("NAs in Pain removed (final dataset) =",dim(cdmd.pvt)[1]," (NAs =",NANs,")"),"\n")
cat(paste("Number of unique zip code =",length(unique(cdmd.pvt$zip))),"\n")

cdmd.pvt$Poverty.q = rep("Low",nrow(cdmd.pvt))
cdmd.pvt$Poverty.q[which(cdmd.pvt$Poverty>quantile(cdmd.pvt$Poverty,p=0.3))] = "Medium"
cdmd.pvt$Poverty.q[which(cdmd.pvt$Poverty>quantile(cdmd.pvt$Poverty,p=0.7))] = "High"
```

```{r}
final.data <- 
  cdmd.pvt %>%
  select("Pain",
         "Poverty","Vacancy",
         "Occupation","Marital.status","Insurance.type",
         "Cancer.stage","Chemo","Surgery",
         "Race","Age",
         "Metastasis","Histology",
         "Primary.Site.1","Primary.Site.2",
         "HER2","ER","PR")
```

# Summary of variables

## Table 1

```{r,echo=FALSE,eval=FALSE}
respvar = "Pain"
explvars = c("Pain","Poverty","Vacancy",
             "Occupation","Marital.status","Insurance.type",
             "Cancer.stage","Chemo","Surgery",
             "Race","Age")

table1(as.formula(paste(" ~ ",paste(c(respvar,explvars),collapse=" + "))), data=final.data)
```

```{r,echo=FALSE,eval=FALSE}
respvar = "Pain"
explvars = c("Pain","Poverty","Vacancy",
             "Occupation","Marital.status","Insurance.type",
             "Cancer.stage","Chemo","Surgery",
             "Race","Age",
             "Metastasis","Histology",
             "Primary.Site.1","Primary.Site.2",
             "HER2","ER","PR")

table1(as.formula(paste(" ~ ",paste(c(respvar,explvars),collapse=" + "))), data=final.data)
```

```{r,echo=FALSE}
explvars = c("Pain","Poverty","Vacancy",
             "Occupation","Marital.status","Insurance.type",
             "Cancer.stage","Chemo","Surgery",
             "Race","Age",
             "Metastasis","Histology",
             "Primary.Site.1","Primary.Site.2",
             "HER2","ER","PR")

get_summary = function(j) {
  x = final.data[,j]
  cname = colnames(final.data)[j]
  if (class(x)=="numeric") {
    tmp.sum = t(t(c(paste0(round(mean(x),digits=2)," (",round(sd(x),digits=2),")"),
                    paste0(round(median(x),digits=2)," [",round(min(x),digits=2),",",round(max(x),digits=2),"]"))))
    rownames(tmp.sum) = paste(cname,c("Mean (SD)","Median [Min,Max]"),sep="_")
  } else {
    tmp.sum = t(t(summary(x)))
    tmp.sum[,1] = paste(tmp.sum[,1],paste0("(",round(100*(tmp.sum[,1]/sum(tmp.sum[,1])),digits=1),"%)"))
    rownames(tmp.sum) = paste(cname,rownames(tmp.sum),sep="_")
  }
  return(tmp.sum)
}
table.sum1 = do.call("rbind",sapply(1:ncol(final.data),get_summary))

get_mean = function(j) {
  x = final.data[,j]
  cname = colnames(final.data)[j]
  formula = as.formula(paste("Pain ~",cname))
  if (class(x)=="numeric") {
    if (! cname == "Pain") {
      fit = glm(formula, data = final.data, family = "poisson")
      tmp.sum = t(t(c(round(exp(coef(summary(fit))[2,1]),digits=2),NA)))
    } else {
      tmp.sum = t(t(c(NA,NA)))
    }
  } else {
    tmp.sum = t(t(sapply(split(final.data[,"Pain"],final.data[,j]),function(x) 
  paste0(round(mean(x),digits=2)," [",paste0(round(quantile(x,c(0.25,0.75)),digits=2),collapse=","),"]"))))
  }
  return(tmp.sum)
}

get_pval = function(j) {
  x = final.data[,j]
  cname = colnames(final.data)[j]
  formula = as.formula(paste("Pain ~",cname))
  if (class(x)=="numeric") {
    if (! cname == "Pain") {
      fit = glm(formula, data = final.data, family = "poisson")
      tmp.sum = t(t(c(transform.p(coef(summary(fit))[2,4],digits=3),NA)))
    } else {
      tmp.sum = t(t(c(NA,NA)))
    }
  } else {
    tmp.sum = t(t(rep(NA,length(levels(x)))))
    pval = compare_means(formula,data=final.data)
    tmp.sum[1,1] = transform.p(min(pval$p[which((! pval$group1 %in% "Unknown") & (! pval$group2 %in% "Unknown"))]),digits=3)
  }
  return(tmp.sum)
}

table.sum1 = cbind(do.call("rbind",sapply(1:ncol(final.data),get_summary)),
                   do.call("rbind",sapply(1:ncol(final.data),get_mean)),
                   do.call("rbind",sapply(1:ncol(final.data),get_pval)))
colnames(table.sum1) = c("No. (%)","Pain, mean [Q1,Q3]","P-value")

## htmlTable
library(htmlTable)

rowlabels = sapply(rownames(table.sum1),function(x) unlist(strsplit(x,split="_"))[1])
htmlTable(table.sum1,
          rgroup = c(gsub(pattern="[.]",replacement=" ",x=explvars[1:length(explvars)])),
          rnames = sapply(rownames(table.sum1),function(x) unlist(strsplit(x,split="_"))[2]),
          n.rgroup = sapply(explvars, function(x) length(which(rowlabels==x))),
          caption = "Table 1. Baseline characteristics, mean and interquartile range of pain intensity, nonparametric rank test")

```

```{r,echo=FALSE,eval=FALSE}
library(stats)
j = 4
colnames(final.data)[j]
kruskal.test(final.data[,"Pain"] ~ final.data[,j])
kruskal.test(final.data[-which(final.data[,j]=="Unknown"),"Pain"] ~ final.data[-which(final.data[,j]=="Unknown"),j])
å
kruskal.test(final.data[which(final.data[,j] %in% c("Employed","Unemployed")),"Pain"] ~ final.data[which(final.data[,j] %in% c("Employed","Unemployed")),j])

wilcox.test(final.data[-which(final.data[,j]=="Unknown"),"Pain"] ~ final.data[-which(final.data[,j]=="Unknown"),j])

pval = compare_means(Pain ~ Occupation,data=final.data)
min(pval$p[which((! pval$group1 %in% "Unknown") & (! pval$group2 %in% "Unknown"))])
```

## Figure 2

**_Pain variable (Pain)_**:

* For each patient, multiple pains were reported at different dates after their initial diagnosis dates. As a representative value of these multiple observations, average was computed for each patient within the 365-day window.

```{r,echo=FALSE,fig.width=8}
# pdf(6,6,file=paste("Pain_hist.pdf"))
barplot(table(cdmd.pvt$Pain),main=paste("Pain intensity"),ylab="Frequency")
# legend("topright",bty="n",legend=c(paste("Mean   =",round(mean(cdmd.pvt$Pain),digits=2)),
                                   # paste("Median =",round(median(cdmd.pvt$Pain),digits=2))))
# dev.off()

```

**_Poverty variable_**:

* Poverty was measured at zip code level. Since we have 34 different zip code, Poverty variable also has 34 unique values each of which corresponds to each zip code.

```{r,echo=FALSE,fig.width=8}

hist(cdmd.pvt$Poverty,main="Poverty")
cat(paste("Number of zipcode =",length(unique(cdmd.pvt$zip))),"\n")
cat(paste("Median of Poverty =",round(median(cdmd.pvt$Poverty),digits=3)),"\n")

```

**_Vacancy variable_**:

```{r,echo=FALSE,fig.width=8}

hist(cdmd.pvt$Vacancy,main="Vacancy")
cat(paste("Number of zipcode =",length(unique(cdmd.pvt$zip))),"\n")
cat(paste("Median of Vacancy =",round(median(cdmd.pvt$Vacancy),digits=3)),"\n")

```

### Poverty

* Pain ~ Poverty

```{r,echo=FALSE,eval=FALSE}
p.glm = glm(Pain ~ Poverty, data = final.data, family = "poisson")
# summary(p.glm)
fhat = function(x) {
  exp(p.glm$coefficients[1]+p.glm$coefficients[2]*x)
}

meandata = data.frame(Poverty = aggregate(Poverty ~ zip,cdmd.pvt,mean)$Poverty,
                      Pain = aggregate(Pain ~ zip,cdmd.pvt,mean)$Pain,
                      zip = aggregate(Poverty ~ zip,cdmd.pvt,mean)$zip)
meandata$size = sapply(meandata$zip,function(x) length(which(cdmd.pvt$zip==x)))
# hist(round(meandata$Pain))
# summary(glm(round(meandata$Pain) ~ meandata$Poverty,family = "poisson"))
# summary(glm.nb(round(meandata$Pain) ~ meandata$Poverty))

# par(mfrow=c(1,2))
cex = exp(2*(meandata$size/(max(meandata$size) - min(meandata$size) + 1)))

pdf(6,6,file=paste("Pain_poverty_plot.pdf"))
par(mar=c(4,4,2,2))
plot(x=meandata$Poverty,y=meandata$Pain,
     xlim=c(0,70),ylim=c(0,6.5),col=brewer.pal(n=11,"BrBG")[9],
     axes=F,
     xlab=NA,ylab=NA,
     cex=cex,lwd=2.5)
box()
abline(0, 0, lty=1, col="grey")
abline(v=0, lty=1, col="grey")
# axis(side=1, tck=-.015,labels=NA)
# axis(side=1, lwd=0, line=-1, cex=0.2,cex.axis=0.9)
axis(side=1, tck=-0.015,lwd=0,line=-.5,cex.axis=1.5)
axis(side=2, tck=-0.015,lwd=0,line=-.8,cex.axis=1.5)
mtext(side=1, "Poverty", line=2.5, cex=1.5)
mtext(side=2, "Pain average", line=2, cex=1.5)

# points(x=meandata$Poverty,y=meandata$Pain,
#        cex=cex,lwd=1.5)
x=seq(0,70,by=1)
points(x=x,y=fhat(x),type="l",lwd=3,
       col=brewer.pal(n=9,"YlOrRd")[7])
dev.off()

# par(mfrow=c(3,3))
pdf(6,6,file=paste("Pain_poverty_legend.pdf"))
par(mar=c(4,4,2,2))
plot(x=c(0,1),y=c(0,1),axes=F,cex=0,xlab=NA,ylab=NA)
# cbind(meandata$size,cex)

x = c(20,40,60,80)
y = c(1.5,2.5,4,6)
points(x=rep(0.4,length(x)),
       y=seq(0.8,0.4,length.out=length(x)),
       cex=y,lwd=2.5,col=brewer.pal(n=11,"BrBG")[9])
text(x=rep(0.55,length(x)),
     y=seq(0.8,0.4,length.out=length(x)),
     labels=x,cex=1.5)
text(x=0.5,y=0.95,
     labels="Number of patients",cex=1.5)
dev.off()

```

### Vacancy

* Pain ~ Vacancy

```{r,echo=FALSE,eval=FALSE}
p.glm = glm(Pain ~ Vacancy, data = final.data, family = "poisson")
# summary(p.glm)
fhat = function(x) {
  exp(p.glm$coefficients[1]+p.glm$coefficients[2]*x)
}

meandata = data.frame(Vacancy = aggregate(Vacancy ~ zip,cdmd.pvt,mean)$Vacancy,
                      Pain = aggregate(Pain ~ zip,cdmd.pvt,mean)$Pain,
                      zip = aggregate(Vacancy ~ zip,cdmd.pvt,mean)$zip)
meandata$size = sapply(meandata$zip,function(x) length(which(cdmd.pvt$zip==x)))
# hist(round(meandata$Pain))
# summary(glm(round(meandata$Pain) ~ meandata$Vacancy,family = "poisson"))
# summary(glm.nb(round(meandata$Pain) ~ meandata$Vacancy))

# par(mfrow=c(1,2))
cex = exp(2*(meandata$size/(max(meandata$size) - min(meandata$size) + 1)))

pdf(6,6,file=paste("Pain_Vacancy_plot.pdf"))
par(mar=c(4,4,2,2))
plot(x=meandata$Vacancy,y=meandata$Pain,
     xlim=c(0,60),ylim=c(0,6.5),col=brewer.pal(n=11,"BrBG")[9],
     axes=F,
     xlab=NA,ylab=NA,
     cex=cex,lwd=2.5)
box()
abline(0, 0, lty=1, col="grey")
abline(v=0, lty=1, col="grey")
# axis(side=1, tck=-.015,labels=NA)
# axis(side=1, lwd=0, line=-1, cex=0.2,cex.axis=0.9)
axis(side=1, tck=-0.015,lwd=0,line=-.5,cex.axis=1.5)
axis(side=2, tck=-0.015,lwd=0,line=-.8,cex.axis=1.5)
mtext(side=1, "Vacancy", line=2.5, cex=1.5)
mtext(side=2, "Pain average", line=2, cex=1.5)

x=seq(0,53,by=1)
points(x=x,y=fhat(x),type="l",lwd=3,
       col=brewer.pal(n=9,"YlOrRd")[7])
dev.off()

```

```{r,echo=FALSE,eval=FALSE}
p.glm = glm(Pain ~ Vacancy, data = cdmd.pvt, family = "poisson")
# summary(p.glm)
# summary(zeroinfl(Pain ~ Poverty, data = cdmd.pvt, dist="poisson"))
fhat = function(x) {
  exp(p.glm$coefficients[1]+p.glm$coefficients[2]*x)
}

meandata = data.frame(Vacancy = aggregate(Vacancy ~ zip,cdmd.pvt,mean)$Vacancy,
                      Pain = aggregate(Pain ~ zip,cdmd.pvt,mean)$Pain,
                      zip = aggregate(Vacancy ~ zip,cdmd.pvt,mean)$zip)
meandata$size = sapply(meandata$zip,function(x) length(which(cdmd.pvt$zip==x)))
# hist(round(meandata$Pain))
# summary(glm(round(meandata$Pain) ~ meandata$Vacancy,family = "poisson"))
# summary(glm.nb(round(meandata$Pain) ~ meandata$Vacancy))

par(mfrow=c(1,2))
cex = exp(2*(meandata$size/(max(meandata$size) - min(meandata$size) + 1)))
# pdf(6,6,file=paste("Pain_Vacancy_plot.pdf"))
plot(x=meandata$Vacancy,y=meandata$Pain,cex=cex,xlab="Vacancy",ylab="Pain average")
# plot(x=aggregate(Vacancy ~ zip,cdmd.pvt,mean)$Vacancy,
#      y=aggregate(Pain ~ zip,cdmd.pvt,mean)$Pain,
#      xlab="Vacancy",ylab=paste(pain.type,"Pain mean within Zip"))
x=seq(0,80,by=1)
points(x=x,y=fhat(x),type="l",lwd=2)
# dev.off()
```

# Analysis

## Figure 2

# Significance

```{r,echo=FALSE,eval=FALSE}
xlist = explvars[4:length(explvars)]

x = xlist[1]
p <- ggboxplot(final.data, x=x, y="Pain",remove="Unknown",
               color = x, palette="jco")
p + stat_compare_means()

x = xlist[2]
p <- ggboxplot(cdmd.pvt, x=x, y="Pain",remove="Unknown",
               color = x, palette="jco")
p + stat_compare_means()

x = xlist[3]
p <- ggboxplot(cdmd.pvt, x=x, y="Pain",remove="Unknown",
               color = x, palette="jco")
p + stat_compare_means()

x = xlist[4]
p <- ggboxplot(cdmd.pvt, x=x, y="Pain",
               color = x, palette="jco")
p + stat_compare_means()

x = xlist[5]
p <- ggboxplot(cdmd.pvt, x=x, y="Pain",
               color = x, palette="jco")
p + stat_compare_means()

x = xlist[6]
p <- ggboxplot(cdmd.pvt, x=x, y="Pain",
               color = x, palette="jco")
p + stat_compare_means()

x = xlist[7]
p <- ggboxplot(cdmd.pvt, x=x, y="Pain",
               color = x, palette="jco")
p + stat_compare_means()

x = xlist[8]
p <- ggboxplot(cdmd.pvt, x=x, y="Pain",
               color = x, palette="jco")
p + stat_compare_means()

x = xlist[9]
p <- ggboxplot(cdmd.pvt, x=x, y="Pain",
               color = x, palette="jco")
p + stat_compare_means()

```

```{r,echo=FALSE,eval=FALSE}
## Figure code - boxplot
xlist = explvars[4:length(explvars)]

# x = "Poverty"
# tmp = cdmd.pvt[,c("Pain","Poverty.f")]
# colnames(tmp) = c("Pain",x)
# tmp$Poverty = factor(tmp$Poverty)
# tmp$Poverty = relevel(tmp$Poverty,"low")
# base <- ggboxplot(tmp, x=x, y="Pain",
#                   color = x, palette="jco",legend="none") +
#   labs(x = NULL, y = "Pain", title = x) + 
#   font("title",size=20,color="black") + 
#   font("x.text", size = 20, color = "black") +
#   font("ylab", size = 20, color = "black") + 
#   font("y.text", size = 20, color = "black") +
#   stat_compare_means(size=7, aes(label = "p.format"),show.legend=F,hide.ns=T,
#                      label.y = 10.5)
# p <- base +
#   theme(plot.title=element_text(face="bold",size=20,margin=margin(t=10,b=20)))
# ggsave(p, file=paste0(x,".pdf"),width=6, height=6)

x = xlist[4]
base <- ggboxplot(final.data, x=x, y="Pain",remove="Unknown",
                  color = x, palette="jco",legend="none") +
  labs(x = NULL, y = "Pain", title = x) + 
  font("title",size=20,color="black") + 
  font("x.text", size = 20, color = "black") +
  font("ylab", size = 20, color = "black") + 
  font("y.text", size = 20, color = "black") +
  stat_compare_means(size=7, aes(label = "p.format"),show.legend=F,hide.ns=T,
                     label.y = 10.5)
p <- base +
  theme(plot.title=element_text(face="bold",size=20,margin=margin(t=10,b=20)))
p
# ggsave(p, file=paste0(x,".pdf"),width=6, height=6)

x = xlist[3]
base <- ggboxplot(final.data, x=x, y="Pain",remove="Unknown",
                  color = x, palette="jco",legend="none") +
  labs(x = NULL, y = "Pain", title = x) + 
  font("title",size=20,color="black") + 
  font("x.text", size = 20, color = "black") +
  font("ylab", size = 20, color = "black") + 
  font("y.text", size = 20, color = "black") +
  stat_compare_means(size=7, aes(label = "p.format"),show.legend=F,hide.ns=T,
                     label.y=10.5)
p <- base +
  theme(plot.title=element_text(face="bold",size=20,margin=margin(t=10,b=20)))
p
# ggsave(p, file=paste0(x,".pdf"),width=6, height=6)

x = "Occupation"
my_comparisons <- list( c("Employed", "Retired"), c("Retired", "Unemployed"), c("Employed", "Unemployed") )
base <- ggboxplot(final.data, x=x, y="Pain",remove="Unknown",
                  color = x, palette="jco",legend="none") +
  labs(x = NULL, y = "Pain", title = x) + 
  font("title",size=20,color="black") + 
  font("x.text", size = 20, color = "black") +
  font("ylab", size = 20, color = "black") + 
  font("y.text", size = 20, color = "black") +
  stat_compare_means(size=7, aes(label = "p.format"),show.legend=F,hide.ns=T,
                     label.y = 15) +
  stat_compare_means(comparisons = my_comparisons, aes(label = ..p.signif..),hide.ns=T)
p <- base +
  theme(plot.title=element_text(face="bold",size=20,margin=margin(t=10,b=20)))
p
# ggsave(p, file=paste0(x,".pdf"),width=6, height=6)

x = "Race"
tmp <-
  cdmd.pvt %>%
  filter(race.f %in% c("White","Black")) %>%
  select("Pain","race.f")
colnames(tmp) = c("Pain",x)
base <- ggboxplot(tmp, x=x, y="Pain",
                  color = x, palette="jco",legend="none") +
  labs(x = NULL, y = "Pain", title = x) + 
  font("title",size=20,color="black") + 
  font("x.text", size = 20, color = "black") +
  font("ylab", size = 20, color = "black") + 
  font("y.text", size = 20, color = "black") +
  stat_compare_means(size=7, aes(label = "p.format"),show.legend=F,hide.ns=T,
                     label.y = 10.5)
p <- base +
  theme(plot.title=element_text(face="bold",size=20,margin=margin(t=10,b=20)))
ggsave(p, file=paste0(x,".pdf"),width=6, height=6)

x = "Age"
tmp <-
  cdmd.pvt %>%
  select("Pain","age.f")
colnames(tmp) = c("Pain",x)
tmp$Age[which(tmp$Age==0)] = "low"
tmp$Age[which(tmp$Age==1)] = "high"
tmp$Age = factor(tmp$Age)
tmp$Age = relevel(tmp$Age,"low")
base <- ggboxplot(tmp, x=x, y="Pain",
                  color = x, palette="jco",legend="none") +
  labs(x = NULL, y = "Pain", title = x) + 
  font("title",size=20,color="black") + 
  font("x.text", size = 20, color = "black") +
  font("ylab", size = 20, color = "black") + 
  font("y.text", size = 20, color = "black") +
  stat_compare_means(size=7, aes(label = "p.format"),show.legend=F,hide.ns=T,
                     label.y = 10.5)
p <- base +
  theme(plot.title=element_text(face="bold",size=20,margin=margin(t=10,b=20)))
ggsave(p, file=paste0(x,".pdf"),width=6, height=6)

x = "Surgery"
tmp <-
  cdmd.pvt %>%
  select("Pain","surgery.f")
colnames(tmp) = c("Pain",x)
base <- ggboxplot(tmp, x=x, y="Pain",
                  color = x, palette="jco",legend="none") +
  labs(x = NULL, y = "Pain", title = x) + 
  font("title",size=20,color="black") + 
  font("x.text", size = 20, color = "black") +
  font("ylab", size = 20, color = "black") + 
  font("y.text", size = 20, color = "black") +
  stat_compare_means(size=7, aes(label = "p.format"),show.legend=F,hide.ns=T,
                     label.y = 10.5)
p <- base +
  theme(plot.title=element_text(face="bold",size=20,margin=margin(t=10,b=20)))
ggsave(p, file=paste0(x,".pdf"),width=6, height=6)

x = "Chemo"
tmp <-
  cdmd.pvt %>%
  select("Pain","chemo.f")
colnames(tmp) = c("Pain",x)
base <- ggboxplot(tmp, x=x, y="Pain",
                  color = x, palette="jco",legend="none") +
  labs(x = NULL, y = "Pain", title = x) + 
  font("title",size=20,color="black") + 
  font("x.text", size = 20, color = "black") +
  font("ylab", size = 20, color = "black") + 
  font("y.text", size = 20, color = "black") +
  stat_compare_means(size=7, aes(label = "p.format"),show.legend=F,hide.ns=T,
                     label.y = 10.5)
p <- base +
  theme(plot.title=element_text(face="bold",size=20,margin=margin(t=10,b=20)))
ggsave(p, file=paste0(x,".pdf"),width=6, height=6)

x = "Stage"
tmp <-
  cdmd.pvt %>%
  select("Pain","stage.f")
colnames(tmp) = c("Pain",x)
my_comparisons <- list( c("1", "2"), c("2", "3"), c("1", "3") )
base <- ggboxplot(tmp, x=x, y="Pain",
                  color = x, palette="jco",legend="none") +
  labs(x = NULL, y = "Pain", title = x) + 
  font("title",size=20,color="black") + 
  font("x.text", size = 20, color = "black") +
  font("ylab", size = 20, color = "black") + 
  font("y.text", size = 20, color = "black") +
  stat_compare_means(size=7, aes(label = "p.format"),show.legend=F,hide.ns=T,
                     label.y = 15) +
  stat_compare_means(comparisons = my_comparisons, aes(label = ..p.signif..),hide.ns=T)
p <- base +
  theme(plot.title=element_text(face="bold",size=20,margin=margin(t=10,b=20)))
ggsave(p, file=paste0(x,".pdf"),width=6, height=6)
```

## Table 2

```{r,echo=FALSE}
respvar = "Pain"

explvars.1 = c("Poverty","Occupation","Marital.status","Insurance.type",
               "Cancer.stage","Chemo","Surgery",
               "Race","Age")
explvars.2 = c("Vacancy","Occupation","Marital.status","Insurance.type",
               "Cancer.stage","Chemo","Surgery",
               "Race","Age")
explvars = c("Poverty","Vacancy",
             "Occupation","Marital.status","Insurance.type",
             "Cancer.stage","Chemo","Surgery",
             "Race","Age")
full.formula.1 = as.formula(paste(respvar," ~ ",paste(explvars.1,collapse=" + ")))
full.formula.2 = as.formula(paste(respvar," ~ ",paste(explvars.2,collapse=" + ")))
full.formula = as.formula(paste(respvar," ~ ",paste(explvars,collapse=" + ")))
```

```{r,echo=FALSE}

## Get the ordered explanatory variables in full model
full.model = zeroinfl(full.formula, data = final.data, dist="poisson")
full.summary = round(cbind(exp(coef(full.model)),exp(confint(full.model)),do.call("rbind",coef(summary(full.model)))[,c(4)]),digits=3)
table.explvars = rownames(full.summary)

final.1 = zeroinfl(full.formula.1, data = final.data, dist="poisson")
final.2 = zeroinfl(full.formula.2, data = final.data, dist="poisson")

# tab_model(final, digits = 3,
#           title = "Final",
#           show.zeroinf=TRUE,show.intercept=TRUE,
#           show.se=TRUE, show.ci=FALSE,collapse.se=FALSE, 
#           show.p = TRUE, p.style="numeric_stars")

```

```{r,echo=FALSE}
# exp(confint(final)) # confidence interval
final.adj.1 = round(cbind(exp(coef(final.1)),exp(confint(final.1)),do.call("rbind",coef(summary(final.1)))[,c(4)]),digits=3)
colnames(final.adj.1)[c(1,4)] = c("Estimate","P")
final.adj.2 = round(cbind(exp(coef(final.2)),exp(confint(final.2)),do.call("rbind",coef(summary(final.2)))[,c(4)]),digits=3)
colnames(final.adj.2)[c(1,4)] = c("Estimate","P")

## Final model (Unadjusted)
final.unadj.list = as.list(NULL)
formula.base = Pain ~ Poverty
for (i in 1:length(explvars)) {
  formula.new = as.formula(paste(sub("Poverty",replacement=explvars[i],as.character(formula.base))[c(2,3)],collapse=" ~ "))
  model.unadj = zeroinfl(formula.new, data = final.data, dist="poisson")
  summary.unadj = round(cbind(exp(coef(model.unadj)),exp(confint(model.unadj)),do.call("rbind",coef(summary(model.unadj)))[,c(4)]),digits=3)
  colnames(summary.unadj)[c(1,4)] = c("Estimate","P")
  colnames(summary.unadj) = paste("Unadj",colnames(summary.unadj),sep=".")
  final.unadj.list[[i]] <- summary.unadj
  rm(summary.unadj)
}
final.unadj = do.call("rbind",final.unadj.list)
final.unadj = final.unadj[which(grepl(pattern="Intercept",rownames(final.unadj))==F),]

## Merge
final.adj.1 = as.data.frame(final.adj.1)
final.adj.1$vars = rownames(final.adj.1)
final.adj.2 = as.data.frame(final.adj.2)
final.adj.2$vars = rownames(final.adj.2)
final.unadj = as.data.frame(final.unadj)
final.unadj$vars = rownames(final.unadj)

final.table0 = merge(final.adj.1,final.adj.2,by="vars",all=T)
final.table1 = merge(final.table0,final.unadj,by="vars",all=T)
# which(! table.explvars %in% final.table0$vars)
# which(! final.table0$vars %in% table.explvars)
final.table = final.table1[match(table.explvars,final.table1$vars),c(10:13,2:9)]
rownames(final.table) = table.explvars

final.table$Unadj.CI = paste(final.table$`Unadj.2.5 %`,final.table$`Unadj.97.5 %`,sep="-")
final.table$CI.x = paste(final.table$`2.5 %.x`,final.table$`97.5 %.x`,sep="-")
final.table$CI.y = paste(final.table$`2.5 %.y`,final.table$`97.5 %.y`,sep="-")
final.table <- final.table %>%
  select(Unadj.Estimate,Unadj.CI,Unadj.P,Estimate.x,CI.x,P.x,Estimate.y,CI.y,P.y)

transform.p = function(x) {
  y = as.character(x)
  y[which(y==0)] = "<.001"
  return(y)
}
final.table$Unadj.P = transform.p(final.table$Unadj.P)
final.table$P.x = transform.p(final.table$P.x)
final.table$P.y = transform.p(final.table$P.y)
```

## Poisson count model
```{r,echo=FALSE}
## tables for publication
library(htmlTable)
# library(Gmisc)
vars.list = sapply(explvars[-c(1,2)],function(x) levels(final.data[,which(colnames(final.data)==x)]))
vars.list = lapply(vars.list,function(x) {x[1]<-paste(x[1],"(reference)"); return(x)})
tablevars = c("(Intercept)","Poverty","Vacancy",unlist(vars.list)) # all variables that will be printed in the table

table.count = final.table[which(grepl("count",rownames(final.table))),]
rownames(table.count) = sapply(rownames(table.count),function(x) unlist(strsplit(x,split="count_"))[2]) # extract count model

table.count <-
  table.count %>%
  add_row(,.before=which(grepl(pattern="Retired",rownames(table.count)))) %>%
  add_row(,.before=1+which(grepl(pattern="Single",rownames(table.count)))) %>%
  add_row(,.before=2+which(grepl(pattern="Medicaid",rownames(table.count)))) %>%
  add_row(,.before=3+which(grepl(pattern="stage2",rownames(table.count)))) %>%
  add_row(,.before=4+which(grepl(pattern="Chemo",rownames(table.count)))) %>%
  add_row(,.before=5+which(grepl(pattern="Surgery",rownames(table.count)))) %>%
  add_row(,.before=6+which(grepl(pattern="Black",rownames(table.count)))) %>%
  add_row(,.before=7+which(grepl(pattern="Age",rownames(table.count))))
colnames(table.count) = c("Estimate","95% CI","P-value",
                          "Estimate","95% CI","P-value",
                          "Estimate","95% CI","P-value")

htmlTable(table.count,
          rgroup = c("","","",gsub(pattern="[.]",replacement=" ",x=explvars[3:length(explvars)])),
          rnames = tablevars,
          n.rgroup = c(1,1,1,sapply(vars.list,length)),
          cgroup = c("Unadjusted","Adjusted","Adjusted"),
          n.cgroup = c(3,3,3))
# library(kableExtra)
# kable(table.count) %>%
#   kable_styling("striped") %>%
#   add_header_above(c(" " = 1, "Unadjusted" = 3, "Adjusted" = 3))

```

## Inflation model
```{r,echo=FALSE}
table.zero = final.table[which(grepl("zero",rownames(final.table))),]
rownames(table.zero) = sapply(rownames(table.zero),function(x) unlist(strsplit(x,split="zero_"))[2]) # extract zero model

table.zero <-
  table.zero %>%
  add_row(,.before=which(grepl(pattern="Retired",rownames(table.zero)))) %>%
  add_row(,.before=1+which(grepl(pattern="Single",rownames(table.zero)))) %>%
  add_row(,.before=2+which(grepl(pattern="Medicaid",rownames(table.zero)))) %>%
  add_row(,.before=3+which(grepl(pattern="stage2",rownames(table.zero)))) %>%
  add_row(,.before=4+which(grepl(pattern="Chemo",rownames(table.zero)))) %>%
  add_row(,.before=5+which(grepl(pattern="Surgery",rownames(table.zero)))) %>%
  add_row(,.before=6+which(grepl(pattern="Black",rownames(table.zero)))) %>%
  add_row(,.before=7+which(grepl(pattern="Age",rownames(table.zero))))
colnames(table.zero) = c("Estimate","95% CI","P-value",
                          "Estimate","95% CI","P-value",
                          "Estimate","95% CI","P-value")

htmlTable(table.zero,
          rgroup = c("","","",gsub(pattern="[.]",replacement=" ",x=explvars[3:length(explvars)])),
          rnames = tablevars,
          n.rgroup = c(1,1,1,sapply(vars.list,length)),
          cgroup = c("Unadjusted","Adjusted","Adjusted"),
          n.cgroup = c(3,3,3))

# kable(table.zero) %>%
#   kable_styling("striped") %>%
#   add_header_above(c(" " = 1, "Unadjusted" = 3, "Adjusted" = 3))
```

# Basic model + more biomarker variables

```{r,echo=FALSE}
respvar = "Pain"

explvars.1 = c("Poverty","Occupation","Marital.status","Insurance.type",
               "Cancer.stage","Chemo","Surgery",
               "Race","Age",
               "Metastasis","Histology",
               "Primary.Site.1","Primary.Site.2",
               "HER2","ER","PR")

explvars.2 = c("Vacancy","Occupation","Marital.status","Insurance.type",
               "Cancer.stage","Chemo","Surgery",
               "Race","Age",
               "Metastasis","Histology",
               "Primary.Site.1","Primary.Site.2",
               "HER2","ER","PR")

explvars = c("Poverty","Vacancy",
             "Occupation","Marital.status","Insurance.type",
             "Cancer.stage","Chemo","Surgery",
             "Race","Age",
             "Metastasis","Histology",
             "Primary.Site.1","Primary.Site.2",
             "HER2","ER","PR")

full.formula.1 = as.formula(paste(respvar," ~ ",paste(explvars.1,collapse=" + ")))
full.formula.2 = as.formula(paste(respvar," ~ ",paste(explvars.2,collapse=" + ")))
full.formula = as.formula(paste(respvar," ~ ",paste(explvars,collapse=" + ")))
```

```{r,echo=FALSE}
## Get the ordered explanatory variables in full model
full.model = zeroinfl(full.formula, data = final.data, dist="poisson")
full.summary = round(cbind(exp(coef(full.model)),exp(confint(full.model)),do.call("rbind",coef(summary(full.model)))[,c(4)]),digits=3)
table.explvars = rownames(full.summary)

final.1 = zeroinfl(full.formula.1, data = final.data, dist="poisson")
final.2 = zeroinfl(full.formula.2, data = final.data, dist="poisson")

# tab_model(final, digits = 3,
#           title = "Final",
#           show.zeroinf=TRUE,show.intercept=TRUE,
#           show.se=TRUE, show.ci=FALSE,collapse.se=FALSE, 
#           show.p = TRUE, p.style="numeric_stars")

```

```{r,echo=FALSE}
# exp(confint(final)) # confidence interval
final.adj.1 = round(cbind(exp(coef(final.1)),exp(confint(final.1)),do.call("rbind",coef(summary(final.1)))[,c(4)]),digits=3)
colnames(final.adj.1)[c(1,4)] = c("Estimate","P")
final.adj.2 = round(cbind(exp(coef(final.2)),exp(confint(final.2)),do.call("rbind",coef(summary(final.2)))[,c(4)]),digits=3)
colnames(final.adj.2)[c(1,4)] = c("Estimate","P")

## Final model (Unadjusted)
final.unadj.list = as.list(NULL)
formula.base = Pain ~ Poverty
for (i in 1:length(explvars)) {
  formula.new = as.formula(paste(sub("Poverty",replacement=explvars[i],as.character(formula.base))[c(2,3)],collapse=" ~ "))
  model.unadj = zeroinfl(formula.new, data = final.data, dist="poisson")
  summary.unadj = round(cbind(exp(coef(model.unadj)),exp(confint(model.unadj)),do.call("rbind",coef(summary(model.unadj)))[,c(4)]),digits=3)
  colnames(summary.unadj)[c(1,4)] = c("Estimate","P")
  colnames(summary.unadj) = paste("Unadj",colnames(summary.unadj),sep=".")
  final.unadj.list[[i]] <- summary.unadj
  rm(summary.unadj)
}
final.unadj = do.call("rbind",final.unadj.list)
final.unadj = final.unadj[which(grepl(pattern="Intercept",rownames(final.unadj))==F),]

## Merge
final.adj.1 = as.data.frame(final.adj.1)
final.adj.1$vars = rownames(final.adj.1)
final.adj.2 = as.data.frame(final.adj.2)
final.adj.2$vars = rownames(final.adj.2)
final.unadj = as.data.frame(final.unadj)
final.unadj$vars = rownames(final.unadj)

final.table0 = merge(final.adj.1,final.adj.2,by="vars",all=T)
final.table1 = merge(final.table0,final.unadj,by="vars",all=T)
# which(! table.explvars %in% final.table0$vars)
# which(! final.table0$vars %in% table.explvars)
final.table = final.table1[match(table.explvars,final.table1$vars),c(10:13,2:9)]
rownames(final.table) = table.explvars

final.table$Unadj.CI = paste(final.table$`Unadj.2.5 %`,final.table$`Unadj.97.5 %`,sep="-")
final.table$CI.x = paste(final.table$`2.5 %.x`,final.table$`97.5 %.x`,sep="-")
final.table$CI.y = paste(final.table$`2.5 %.y`,final.table$`97.5 %.y`,sep="-")
final.table <- final.table %>%
  select(Unadj.Estimate,Unadj.CI,Unadj.P,Estimate.x,CI.x,P.x,Estimate.y,CI.y,P.y)

final.table$Unadj.P = transform.p(final.table$Unadj.P)
final.table$P.x = transform.p(final.table$P.x)
final.table$P.y = transform.p(final.table$P.y)
```

## Poisson count model:
```{r,echo=FALSE}
## tables for publication
library(htmlTable)
# library(Gmisc)
vars.list = sapply(explvars[-c(1,2)],function(x) levels(final.data[,which(colnames(final.data)==x)]))
vars.list = lapply(vars.list,function(x) {x[1]<-paste(x[1],"(reference)"); return(x)})
tablevars = c("(Intercept)","Poverty","Vacancy",unlist(vars.list)) # all variables that will be printed in the table

table.count = final.table[which(grepl("count",rownames(final.table))),]
rownames(table.count) = sapply(rownames(table.count),function(x) unlist(strsplit(x,split="count_"))[2]) # extract count model

table.count <-
  table.count %>%
  add_row(,.before=which(grepl(pattern="Retired",rownames(table.count)))) %>%
  add_row(,.before=1+which(grepl(pattern="Single",rownames(table.count)))) %>%
  add_row(,.before=2+which(grepl(pattern="Medicaid",rownames(table.count)))) %>%
  add_row(,.before=3+which(grepl(pattern="stage2",rownames(table.count)))) %>%
  add_row(,.before=4+which(grepl(pattern="Chemo",rownames(table.count)))) %>%
  add_row(,.before=5+which(grepl(pattern="Surgery",rownames(table.count)))) %>%
  add_row(,.before=6+which(grepl(pattern="Black",rownames(table.count)))) %>%
  add_row(,.before=7+which(grepl(pattern="Age",rownames(table.count)))) %>%
  add_row(,.before=8+which(grepl(pattern="Metastasis",rownames(table.count)))) %>%
  add_row(,.before=9+which(grepl(pattern="lobular",rownames(table.count)))) %>%
  add_row(,.before=10+which(grepl(pattern="Central",rownames(table.count)))) %>%
  add_row(,.before=11+which(grepl(pattern="Midline",rownames(table.count)))) %>%
  add_row(,.before=12+which(grepl(pattern="HER2Positive",rownames(table.count)))) %>%
  add_row(,.before=13+which(grepl(pattern="ERPositive",rownames(table.count)))) %>%
  add_row(,.before=14+which(grepl(pattern="PRPositive",rownames(table.count))))
colnames(table.count) = c("Estimate","95% CI","P-value",
                          "Estimate","95% CI","P-value",
                          "Estimate","95% CI","P-value")

htmlTable(table.count,
          rgroup = c("","","",gsub(pattern="[.]",replacement=" ",x=explvars[3:length(explvars)])),
          rnames = tablevars,
          n.rgroup = c(1,1,1,sapply(vars.list,length)),
          cgroup = c("Unadjusted","Adjusted","Adjusted"),
          n.cgroup = c(3,3,3))
# library(kableExtra)
# kable(table.count) %>%
#   kable_styling("striped") %>%
#   add_header_above(c(" " = 1, "Unadjusted" = 3, "Adjusted" = 3))

```

## Inflation model:
```{r,echo=FALSE}
table.zero = final.table[which(grepl("zero",rownames(final.table))),]
rownames(table.zero) = sapply(rownames(table.zero),function(x) unlist(strsplit(x,split="zero_"))[2]) # extract zero model

table.zero <-
  table.zero %>%
  add_row(,.before=which(grepl(pattern="Retired",rownames(table.zero)))) %>%
  add_row(,.before=1+which(grepl(pattern="Single",rownames(table.zero)))) %>%
  add_row(,.before=2+which(grepl(pattern="Medicaid",rownames(table.zero)))) %>%
  add_row(,.before=3+which(grepl(pattern="stage2",rownames(table.zero)))) %>%
  add_row(,.before=4+which(grepl(pattern="Chemo",rownames(table.zero)))) %>%
  add_row(,.before=5+which(grepl(pattern="Surgery",rownames(table.zero)))) %>%
  add_row(,.before=6+which(grepl(pattern="Black",rownames(table.zero)))) %>%
  add_row(,.before=7+which(grepl(pattern="Age",rownames(table.zero)))) %>%
  add_row(,.before=8+which(grepl(pattern="Metastasis",rownames(table.zero)))) %>%
  add_row(,.before=9+which(grepl(pattern="lobular",rownames(table.zero)))) %>%
  add_row(,.before=10+which(grepl(pattern="Central",rownames(table.zero)))) %>%
  add_row(,.before=11+which(grepl(pattern="Midline",rownames(table.zero)))) %>%
  add_row(,.before=12+which(grepl(pattern="HER2Positive",rownames(table.zero)))) %>%
  add_row(,.before=13+which(grepl(pattern="ERPositive",rownames(table.zero)))) %>%
  add_row(,.before=14+which(grepl(pattern="PRPositive",rownames(table.zero))))
colnames(table.zero) = c("Estimate","95% CI","P-value",
                          "Estimate","95% CI","P-value",
                          "Estimate","95% CI","P-value")

htmlTable(table.zero,
          rgroup = c("","","",gsub(pattern="[.]",replacement=" ",x=explvars[3:length(explvars)])),
          rnames = tablevars,
          n.rgroup = c(1,1,1,sapply(vars.list,length)),
          cgroup = c("Unadjusted","Adjusted","Adjusted"),
          n.cgroup = c(3,3,3))

# kable(table.zero) %>%
#   kable_styling("striped") %>%
#   add_header_above(c(" " = 1, "Unadjusted" = 3, "Adjusted" = 3))
```